Adding selection to \stdpopsim: oh how the tables have turned

### 2024-12-10 analysis2_manuscript.tex
some revisions (#29)
* fig legends; fix schematic

* changes

* finished results and intro

* methods; + minor discussion  stuff

* update figs (ticks were fed up)

* minor fix

* typo
citation cleanup

@@

    \subsection*{Simulations}
    In this study we performed simulations of the human and vaquita porpoise genomes
    using the \stdpopsim{} framework. As stated above, simulations included a model 
    of (negtive) selection acting on exonic regions, matched to the human and vaquita porpoise genomes.
    Three replicates of whole-genome simulations were performed for each species, with 
    a sample size of 100 diploid individuals. Demographic models are as stated above, each
    using the exon annotations from the human and vaquita porpoise genomes respectively.
    For simulations with selection, we sampled 10 diploid individuals from each of two populations (YRI and CEU). 
    To decrease runtimes, we used a SLiM scaling factor of 2 with a burn-in of 2N generations.

@@

    \subsection*{Selective sweeps detection}
    We ran simulations of human chromosome 1 using the three population out-of-Africa model
    (identified as OutOfAfrica\_3G09 in the stdpopsim library; \cite{gutenkunst_inferring_2009})
    and the genetic map estimated in \citet{frazer_second_2007} (HapMapII\_GRCh38).
    For computational efficiency, we simulated a 5Mb region flanking 100 evenly distributed points along the chromosome 1.
    We generated 1000 replicates of each of three classes of simulations:
    (i) neutral (with the specified demography and genetic map),
    (ii) background selection, in which we applied the DFE estimated in \citet{kim_inference_2017} (Gamma\_K17) to exon annotations (taken from Ensembl; ensembl\_havana\_104\_exons),
    and (iii) sweep with background selection, in which we simulated the hard sweep as described with the addition of constraint in exonic regions (same as for \emph{ii}).
    In the sweep simulations, we placed a single beneficial mutation in the center of the simulated region with a selection coefficient of $s = 0.03$, and only sampled if the beneficial mutation reached terminal frequency of 0.95 or greater.

    We then benchmarked the performance of three methods for detecting sweeps: 1) \sweepfinder \citep{degiorgio2016sweepfinder2},
    2) \diploshic \citep{kern2018diplos}, and 3) an empirical cutoff for nucleotide diversity ($\pi$) to find regions of reduced diversity.
    \diploshic was trained on data simulated using \texttt{discoal} \citep{kern2016discoal}.
    Nucleotide diversity was computed directly from the underlying tree sequences using \tskit \citep{ralph_efficiently_2020} over 10 equally sized and non-overlapping subwindows.
    The composite likelihood ratio (output by \sweepfinder) was computed at 21 uniformly distributed points along the simulated region.
    To train the \diploshic classifier, 
    For training simulations, we simulated 20 haploid genomic regions of 1.21Mbp
    with a constant population size of 10,000 individuals, a mutation rate of 2e-8, and a recombination rate of 2e-8.
    For \texttt{discoal} simulations that included selective sweeps,
    selection coefficients were drawn from a uniform distribution between 0.001 and 0.10.
    The time of the sweep was chosen from a uniform distribution between 0 and 0.01 Ne generations in the past.
    For soft sweeps, the frequency at which selection was introduced was drawn from a uniform distribution between 1e-4 and 0.1.
    All \texttt{discoal} training simulations were done under a model of equilibrium population size, 
    so we expected it to perform suboptimally given the variation in population size in the out-of-Africa model.

>>>>

\newcommand{\tskit}{\texttt{tskit}\xspace}

@@

    \subsection*{Simulations}
    In this study we performed simulations of the human and vaquita porpoise genomes
    using the \stdpopsim{} framework. As stated above, simulations included a model 
    of (negative) selection acting on exonic regions, matched to the human and vaquita porpoise genomes.
    Three replicates of whole-genome simulations were performed for each species, with 
    a sample size of 100 diploid individuals. Demographic models are as stated above, each
    using the exon annotations from the human and vaquita porpoise genomes respectively.
    For simulations with selection, we sampled 10 diploid individuals from each of two populations (YRI and CEU). 
    To decrease runtimes, we used a SLiM scaling factor of 2 with a burn-in of 2N generations.

@@

    \subsection*{Selective sweeps detection}
    We ran simulations of human chromosome 1 using the three population out-of-Africa model
    (identified as OutOfAfrica\_3G09 in the stdpopsim library; \cite{gutenkunst2009inferring})
    and the genetic map estimated in \citet{international2007second} (HapMapII\_GRCh38).
    For computational efficiency, we simulated a 5Mb region flanking 100 evenly distributed points along the chromosome 1.
    We generated 1000 replicates of each of three classes of simulations:
    (i) neutral (with the specified demography and genetic map),
    (ii) background selection, in which we applied the DFE estimated in \citet{kim2017inference} (Gamma\_K17) to exon annotations (taken from Ensembl; ensembl\_havana\_104\_exons),
    and (iii) sweep with background selection, in which we simulated the hard sweep as described with the addition of constraint in exonic regions (same as for \emph{ii}).
    In the sweep simulations, we placed a single beneficial mutation in the center of the simulated region with a selection coefficient of $s = 0.03$, and only sampled if the beneficial mutation reached terminal frequency of 0.95 or greater.

    We then benchmarked the performance of three methods for detecting sweeps: 1) \sweepfinder \citep{degiorgio2016sweepfinder2},
    2) \diploshic \citep{kern2018diplos}, and 3) an empirical cutoff for nucleotide diversity ($\pi$) to find regions of reduced diversity.
    \diploshic was trained on data simulated using \texttt{discoal} \citep{kern2016discoal}.
    Nucleotide diversity was computed directly from the underlying tree sequences using \tskit \citep{ralph2020efficiently} over 10 equally sized and non-overlapping subwindows.
    The composite likelihood ratio (output by \sweepfinder) was computed at 21 uniformly distributed points along the simulated region.
    To train the \diploshic classifier, 
    For training simulations, we simulated 20 haploid genomic regions of 1.21Mbp
    with a constant population size of 10,000 individuals, a mutation rate of 2e-8, and a recombination rate of 2e-8.
    For \texttt{discoal} simulations that included selective sweeps,
    selection coefficients were drawn from a uniform distribution between 0.001 and 0.10.
    The time of the sweep was chosen from a uniform distribution between 0 and 0.01 Ne generations in the past.
    For soft sweeps, the frequency at which selection was introduced was drawn from a uniform distribution between 1e-4 and 0.1.
    All \texttt{discoal} training simulations were done under a model of equilibrium population size, 
    so we expected it to perform suboptimally given the variation in population size in the out-of-Africa model.
